\LARGE \bf Generalizing Informed Sampling for Asymptotically Optimal Sampling-based Planning via Markov Chain Monte Carlo Methods

### 2017-09-14 paper.tex
sync
add level set plot

\author{
Daqing Yi$^{*}$,
Rohan Thakker$^{*}$,
Cole Gulino$^{}$, 
Oren Salzman$^{}$, and
Siddhartha Srinivasa$^{}$% <-this % stops a space
\thanks{*Daqing Yi and R. Thakker contributed equally to this paper.}
\thanks{This work was (partially) funded by the National Science Foundation IIS (\#1409003), Toyota Motor Engineering \& Manufacturing (TEMA), and the Office of Naval Research.}% <-this % stops a space
\thanks{$^{1}${\tt\small \{rthakker, cgulino, daqingy, osalzman, ss5\} @andrew.cmu.edu}}%
%
\\        
Robotics Institute, Carnegie Mellon University$^{1}$
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

\end{abstract}

@@

While our approach can be used with any Markov chain Monte Carlo (MCMC) method, it is especially suited to be used with hit and run~\cite{S84,KSZ11}.
Roughly speaking, this is because hit and run (detailed in Sec~\ref{sec:related_work})
performs a series of one-dimensional rejection samples which are extremely fast to compute, even in high-dimensional spaces.

@@

\begin{figure*}[t!]
	\centering
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[height=3cm]{fig/motivation/slow1}
		\includegraphics[height=3cm]{fig/motivation/slow2}
		\includegraphics[height=3cm]{fig/motivation/slow3}
		\includegraphics[height=3cm]{fig/motivation/slow4}
		\includegraphics[height=3cm]{fig/motivation/slowmotion}
		\caption{Zero goal velocity}
		\label{fig:motivation:slow}
	\end{subfigure}
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[height=3cm]{fig/motivation/fast1}
		\includegraphics[height=3cm]{fig/motivation/fast2}
		\includegraphics[height=3cm]{fig/motivation/fast3}
		\includegraphics[height=3cm]{fig/motivation/fast4}
		\includegraphics[height=3cm]{fig/motivation/fastmotion}
		\caption{Non-zero goal velocity}
		\label{fig:motivation:fast}
	\end{subfigure}
	\caption{HERB moves right arm from a start configuration to a goal configuration, which are very close.
    When the goal velocity is non-zero, HERB needs to move right arm further away to accelerate.
	}
	\label{fig:motivation}
\end{figure*}

\os{structure}

@@

Of specific interest to our work are approaches that produce samples in the informed set~\Cinf.
As mentioned in Sec.~\ref{sec:intro} Gammel et al.~\cite{GSB14} describe an approach to sample uniformly in~\Cinf for the specific case of where $\calX = \R^d$ and when optimizing for path length.
To the best of our knowledge, the only method to produce samples in non-Euclidean spaces that are applied to motion planning problems (other than rejection sampling) is HRS by Kunz et al.~\cite{KTC16}.

@@

One variant, which is of special interest to us is 
hit-and-run~\cite{S84,KSZ11}.
Here, given the current point $x_i$ the next point~$x_{i+1}$ in the Markov chain is produced by sampling a random direction~$\theta$ on the surface of the unit sphere centered at $x_{i+1}$. This defines a ray $r_i$ rooted at $x_i$ and passing through~$\theta$. The point $x_{i+1}$ is chosen by randomly sampling a point on $r_i$.
This algorithm is considered to be one of the most efficient algorithms for generating an asymptotically uniform point if the set under consideration is convex~\cite{L99, LV06}
and it can also be extended to sample points that converge to an arbitrary target distribution in total variation~\cite{BRS93, RS94}.

The attractiveness of hit and run for our problem domain stems from the fact that it performs a series of one-dimensional rejection samples which are extremely fast to compute, even in high-dimensional spaces. 
Finally, it is worth noting that we are not the first to apply hit and run for motion-planning problems.
Recently~\cite{YPVA17} was used as an alternative to RRT to produce \emph{feasible motions} (and not high-quality paths). 
Interestingly the paper concludes with the statement ``\emph{One drawback is that the sample paths for Hit and-Run have no pruning and are therefore longer than the RRT paths. Hybrid approaches that yield short paths but also explore quickly are a promising future direction.}''
Our paper can be seen as a hybrid approach marrying sampling-based planning with MCMC-based approaches.

@@

\subsection{Motivating example---Minimal-Time Double Integrator}
\label{sec:mtdi}

\os{mention~\cite{LM97} here or in related work}

To understand why we resort to optimization-based methods and do not attempt to provide a closed-form solution to sample $\Cinf$ we study the structure of the informed set for a simple yet important dynamical system---the double integrator minimizing time (MTDI). 
Here, we are given a one-dimensional point robot with bounded acceleration moving amid obstacles. We wish to compute the minimal-time trajectory between two states $x_s, x_g$.
A state $x \in \calX$ in this model is defined by 
the position $q \in \R$
and
the velocity $\dot{q}\in \R$ of the robot.
%each coordinate axis.
The system dynamics are described by:
\begin{equation}
\begin{bmatrix}
	\dot{q} \\
	\ddot{q}
\end{bmatrix}
=
\begin{bmatrix}
	0 & 1 \\
	0 & 0
\end{bmatrix}
\begin{bmatrix}
	{q} \\
	\dot{q}
\end{bmatrix}
+
\begin{bmatrix}
	0 \\
	1
\end{bmatrix}
u
\end{equation}
%\begin{equation}
%x = (q, \dot{q}); 
%\hspace{5mm}
%u = \ddot{q}.
%\end{equation}
Here, the control 
%$u \in [u_{\text{min}}, u_{\text{max}}]$ 
$u \in [\underline{u}, \overline{u}]$ 
is the (bounded) acceleration\footnote{One can also bound the velocity of the system. However, to simplify the exposition, we only constrain the system with acceleration bounds.}.

@@

@@

In this section we describe our approach to efficiently produce a sequence of samples in the informed set~\Cinf given a specific cost~$c_{\text{best}}$ of trajectory $x_{best}(t)$.
Our algorithm (see Alg.~\ref{alg:mcmc_informed_sampling}) starts by randomly sampling a state~$x_\text{rand} \in \calX$ (line~2).
This sample is used to produces a 
sample~$x_0 \in \partial\Cinf$ 
%sample~$x_\text{boundary} \in \partial\Cinf$ 
which lies (approximately) on the boundary of the informed set (Sec.~\ref{subsec:grad}) by applying Newton-Raphson method~\cite{RT06} (line~3).
Finally,~$x_0$ 
%use~$x_\text{boundary}$ 
is used (lines~4-6)
to initialize an MCMC algorithm which outputs a series of samples within \Cinf by constructing a Markov Chain.
This is done until some termination criterion is met and the entire process is repeated (Sec.~\ref{subsec:mcmc}). 
%Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}). 
For a visualization of the approach, see Fig.~\ref{fig:alg}.

It is important to note that at each iteration, we will query our algorithm for one sample. However, this sample depends on the previous iterations of the algorithm which is why we describe it as a method to produce a sequence of samples.
Furthermore, the value~$c_{\text{best}}$ can decrease between consecutive iterations.
This will occur if the search algorithm that uses our sampler finds a path to the goal whose cost is lower than~$c_{\text{best}}$.

\os{mention that it is AO}

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\LOOP
			\STATE $x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( method )$
			\WHILE {\texttt{MCMC\_continue\_chain()}}
			\STATE $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$
			\STATE \textbf{output} $x_{i}$
			\ENDWHILE
		\ENDLOOP 
%		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
%%		\STATE $x_{\text{boundary}} 
%		\STATE $x_{0} 
%		 \leftarrow
%		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
%%		\STATE $x_{0} \leftarrow x_{\text{boundary}}$
%		\WHILE {\texttt{MCMC\_continue\_chain()}}
%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$; \hspace{3mm} $i\leftarrow i+1$
%			\STATE \textbf{output} $x_{i}$
%		\ENDWHILE		
%		\FOR {$i = 1$ to $N$}
%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$
%			\STATE \textbf{output} $x_{i}$
%		\ENDFOR
   	\end{algorithmic}
	\caption{MCMC-based Informed Sampling $(c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

\subsection{Sampling the boundary of \Cinf}
\label{subsec:grad}
We can use either of the following two approaches to sample points in the informed space:
\subsubsection{Newton-Raphson Method}
In theory, MCMC methods converge to the desired distribution regardless of the initial sample used to seed the chain.
In our setting, the probability distribution~\Pinf is defined by having all points in \Cinf distributed uniformly
while 
the probability of sampling any configuration $x \in \calX \setminus \Cinf$ is zero.
A common practice to avoid starting biases in MCMC-type algorithm is to discard an initial set of samples (a process referred to as ``burn-in'')~\cite{ADDJ03}.

In our setting, we are only interested in points in~\Cinf, thus we can start the Markov Chain in~\Cinf and avoid this burn-in stage. 
To do so we apply  Newton-Raphson's method to obtain a sample on $\partial \Cinf$. 
However, our cost function is not necessarily continuous (see Sec.~\ref{sec:mtdi}), thus if no sample was produced on $\partial \Cinf$ after a predefined number of steps, we restart the algorithm using a different random state~$x_\text{rand}$
(see also Sec.~\ref{subsec:mcmc}).

\subsubsection{Previous-Trajectory Method}

\subsection{Sampling the interior of \Cinf via MCMC}
\label{subsec:mcmc}

Our approach is general and can be applied to any MCMC algorithm (see Sec.~\ref{sec:related_work}).
We demonstrate how to instantiate it with two different algorithms: 
Hit and Run and Metropolis-Hastings.

For each algorithm, 
we describe how it is used to 
(i)~sample points 
(\texttt{MCMC\_sample, line 5})
and 
(ii)~decide if to continue sampling along the Markov Chain or to restart the entire process 
(\texttt{MCMC\_continue\_chain, line 4}).

\subsubsection{Hit and Run}
\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $d \leftarrow$ \texttt{sample\_random\_direction}$()$
		\STATE $ L(\lambda) = \{  x \mid x = x_{i-1} + \lambda d_i \} $
		\STATE $ \lambda^{+} \leftarrow \sup L(\lambda) $; 
				\hspace{3mm} 
			   $ \lambda^{-} \leftarrow \inf L(\lambda) $

@@

\begin{algorithm}[t]
	\begin{algorithmic}[1]
	\IF{ $method = \texttt{Newton-Rapson}$  }
		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
		\STATE $x_{0} \leftarrow \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
	\ELSIF{ $method = \texttt{Previous-Trajectory}$ }
	    \STATE $t_{rand} \leftarrow \texttt{sample\_random\_time()}$
	    \STATE $x_{0} \leftarrow x_{best}(t_{rand})$
	\ENDIF
	\RETURN $x_0$
	\end{algorithmic}
	\caption{sample\_in\_informed\_space$(method, c_{best})$}
	\label{alg:sample_inf}	
\end{algorithm}

@@

\begin{figure}[t!]
	\centering
	\begin{subfigure}[b]{0.48\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/levelset}
		\caption{2DoF - 4D. TODO:Time per sample vs level set}
		\label{fig:sampling_efficiency:levelset}
	\end{subfigure}
	%\begin{subfigure}[b]{0.48\linewidth}
	%	\includegraphics[width=\linewidth]{fig/sampling_efficiency/levelset_6d}
	%	\caption{6DoF - 12D. TODO:Time per sample vs level set}
	%	\label{fig:sampling_efficiency:6d}
	%\end{subfigure}
	 \\
	\begin{subfigure}[b]{0.78\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{2DoF - 4D TODO: rescale x axis}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
% 	\begin{subfigure}[b]{0.32\textwidth}
% 		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_3d}
% 		\caption{3DoF - 6D}
% 		\label{fig:sampling_efficiency:3d}
% 	\end{subfigure}	
	\begin{subfigure}[b]{0.78\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{6DoF - 12D. TODO: rescale x axis}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Sampling efficiency of different sampling methods}
	\label{fig:sampling_efficiency}
\end{figure} 
vantage of sampling a near state.
Both the per-sampling time of HRS 
It is noticeable that both x axis and y axis are converted by $ \log_{10} () $ in Figure~\ref{fig:sampling_efficiency:2d} and Figure~\ref{fig:sampling_efficiency:6d}.
This reserves the linearity of the data from rejection sampler, but also scales the difference in a smaller range for comparison.
Keep in mind that the differences in unit (ms) are much bigger.
\dy{TODO: change MCMC to MH (Metropolis Hasting) in the plots}
Metropolis-Hasting shows consistent performance when the informed set volume ratio is decreased, because it takes the adand HNR increases with the decrease of the informed set volume ratio. 
When informed set volume ratio is relatively high, it is easy to generate qualified samples.
All the samplers have close performances.
RS and HRS can be the most efficient because of its simple design.
But when problems get harder, HNR outperforms HRS significantly.

@@

\section{Conclusion}
\os{Discuss param tuning, we did nothing + there is a wealth of literature}

\os{use with BIT*, challenge - distance function is not a metric}

The Met

What MCMC algorithm is the best to use?

Parameter tuning

Application to different state spaces

Parallel chains

estimating volume of informed space

>>>>

\author{
Daqing Yi$^{*}$,
Rohan Thakker$^{*}$,
Cole Gulino$^{}$, 
Oren Salzman$^{}$ and
Siddhartha Srinivasa$^{}$% <-this % stops a space
\thanks{*Daqing Yi and R. Thakker contributed equally to this paper.}
\thanks{This work was (partially) funded by the National Science Foundation IIS (\#1409003), Toyota Motor Engineering \& Manufacturing (TEMA), and the Office of Naval Research.}% <-this % stops a space
\thanks{$^{1}${\tt\small \{rthakker, cgulino, daqingy, osalzman, ss5\} @andrew.cmu.edu}}%
%
\\        
Robotics Institute, Carnegie Mellon University$^{1}$
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Asymptotically-optimal motion planners such as RRT* have been shown to incrementally approximate the shortest path between a start and goal states.
Once an initial solution is found, their performance can be dramatically improved by restricting subsequent samples to regions of the state space that can potentially improve the current solution.
When the motion planning problem lies in a Euclidean space, this region~$\Cinf$, called the informed set, can be sampled directly.
However, when planning with differential constraints in non-Euclidean state spaces, no analytic solutions exists to sampling~$\Cinf$ directly.

State-of-the-art approaches to sampling~$\Cinf$ in such domains such as Hierarchical Rejection Sampling (HRS) may still be slow in high-dimensional state space.
This may cause the planning algorithm to spend most of its time trying to produces samples in~$\Cinf$ rather than explore it.
In this paper, we suggest an alternative approach to produce samples in the informed set~\Cinf for a wide range of settings.
Our main insight is to recast this problem as one of sampling uniformly within a level set of an implicit function.
This recasting enables us to apply Monte Carlo sampling methods, used very effectively in the Machine Learning and Optimization communities, to solve our problem.
We show for a wide range of scenarios that using our sampler can accelerate the convergence rate to high-quality solutions in high-dimensional problems.
\end{abstract}

@@

While our approach can be used with any Markov chain Monte Carlo (MCMC) method, it is especially suited to be used with Hit-and-Run~\cite{S84,KSZ11}.
Roughly speaking, this is because Hit-and-Run (detailed in Sec~\ref{sec:related_work})
performs a series of one-dimensional rejection samples which are extremely fast to compute, even in high-dimensional spaces.

@@

\begin{figure*}[t!]
	\centering
	\begin{minipage}[t]{0.64\linewidth}
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[height=2.7cm]{fig/motivation/slow1}
		\includegraphics[height=2.7cm]{fig/motivation/slow2}
		\includegraphics[height=2.7cm]{fig/motivation/slow3}
		\includegraphics[height=2.7cm]{fig/motivation/slow4}
		%\includegraphics[height=3cm]{fig/motivation/slowmotion}
		\caption{Zero goal velocity}
		\label{fig:motivation:slow}
	\end{subfigure}
	\begin{subfigure}[b]{\textwidth}
		\centering
		\includegraphics[height=2.7cm]{fig/motivation/fast1}
		\includegraphics[height=2.7cm]{fig/motivation/fast2}
		\includegraphics[height=2.7cm]{fig/motivation/fast3}
		\includegraphics[height=2.7cm]{fig/motivation/fast4}
		%\includegraphics[height=3cm]{fig/motivation/fastmotion}
		\caption{Non-zero goal velocity}
		\label{fig:motivation:fast}
	\end{subfigure}
    \end{minipage}
    \begin{minipage}[t]{0.33\linewidth}
    	\begin{subfigure}[c]{\textwidth}
    		\includegraphics[width=\textwidth]{fig/motivation/phase_plot}
    		\caption{Non-zero goal velocity}
    		\label{fig:motivation:phase_plot}
        \end{subfigure}
   	\end{minipage}
	\caption{HERB moves right arm from a start configuration to a goal configuration, which are very close.
    When the goal velocity is non-zero, HERB needs to move right arm further away to accelerate.
	}
	\label{fig:motivation}
\end{figure*}

The rest of the paper is structured as follows: after describing related work in Sec.~\ref{sec:related_work}, we formally define our problem in Sec.~\ref{sec:pdef}.
We then provide in Sec.~\ref{sec:motivation} an intuitive description of the challenges faced in sampling within the informed set for our planning domains.
We continue in Sec.~\ref{sec:algorithm} with a description of our algorithm and present experimental evaluations in Sec.~\ref{sec:eval}.
Finally, we conclude with a discussion in Sec.~\ref{sec:future}.

@@

Of specific interest to our work are approaches that produce samples in the informed set~\Cinf.
As mentioned in Sec.~\ref{sec:intro} Gammel et al.~\cite{GSB14} describe an approach to sample uniformly in~\Cinf for the specific case where $\calX = \R^d$ and when optimizing for path length.
To the best of our knowledge, the only method to produce samples in non-Euclidean spaces that can be  applied to motion planning problems (other than rejection sampling) is HRS by Kunz et al.~\cite{KTC16}.

@@

One variant, which is of special interest to us is 
Hit and Run~\cite{S84,KSZ11}.
Here, given the current point $x_i$ the next point~$x_{i+1}$ in the Markov chain is produced by sampling a random direction~$\theta$ on the surface of the unit sphere centered at $x_{i+1}$. This defines a ray $r_i$ rooted at $x_i$ and passing through~$\theta$. The point $x_{i+1}$ is chosen by randomly sampling a point on $r_i$.
This algorithm is considered to be one of the most efficient algorithms for generating an asymptotically uniform point if the set under consideration is convex~\cite{L99, LV06}
and it can also be extended to sample points that converge to an arbitrary target distribution in total variation~\cite{BRS93, RS94}.

The attractiveness of Hit and Run for our problem domain stems from the fact that it performs a series of one-dimensional rejection samples which are extremely fast to compute, even in high-dimensional spaces. 
Finally, it is worth noting that we are not the first to apply Hit and Run for motion-planning problems.
Recently~\cite{YPVA17} was used as an alternative to RRT to produce \emph{feasible motions} (and not high-quality paths). 
Interestingly the paper concludes with the statement ``\emph{One drawback is that the sample paths for Hit and-Run have no pruning and are therefore longer than the RRT paths. Hybrid approaches that yield short paths but also explore quickly are a promising future direction.}''
Our paper can be seen as a hybrid approach marrying sampling-based planning with MCMC-based approaches.

@@

In this work we consider the problem of efficiently producing samples within~$\Cinf$.
These samples will be used within the informed RRT* framework to efficiently and incrementally compute trajectories of decreasing cost, converging to the optimal trajectory.
\os{mention that it is AO}
\begin{prop}
	\dy{claim that once informed sampler provides uniform samplings in any given level set in a non-euclidean state space, the corresponding informed RRT* could asymptotically find an optimal solution.}
\end{prop}

\section{Motivation---\Cinf in kinodynamic state spaces}
\label{sec:mtdi}

In this section we properly motivate this work.
Specifically, we start by describing the differences in between planning in Euclidean configuration spaces (also called geometric planning) nd non-Euclidean  state spaces.

%\os{mention~\cite{LM97} here or in related work}
\subsection{Geometric vs. Kinodynamic planning}

\os{refer to Fig.~2 or the updated one.}

\subsection{Minimal Time Double Integrator}
To understand why we resort to optimization-based methods and do not attempt to provide a closed-form solution to sample $\Cinf$ we study the structure of the informed set for a simple yet important dynamical system---the double integrator minimizing time (MTDI). 
Here, we are given a one-dimensional point robot with bounded acceleration moving amid obstacles. We wish to compute the minimal-time trajectory between two states $x_s, x_g$.
A state $x \in \calX$ in this model is defined by 
the position $q \in \R$
and
the velocity $\dot{q}\in \R$ of the robot.
%each coordinate axis.
The system dynamics are described by:
\begin{equation}
\begin{bmatrix}
	\dot{q} \\
	\ddot{q}
\end{bmatrix}
=
\begin{bmatrix}
	0 & 1 \\
	0 & 0
\end{bmatrix}
\begin{bmatrix}
	{q} \\
	\dot{q}
\end{bmatrix}
+
\begin{bmatrix}
	0 \\
	1
\end{bmatrix}
u.
\end{equation}
%\begin{equation}
%x = (q, \dot{q}); 
%\hspace{5mm}
%u = \ddot{q}.
%\end{equation}
Here, the control 
%$u \in [u_{\text{min}}, u_{\text{max}}]$ 
$u \in [\underline{u}, \overline{u}]$ 
is the (bounded) acceleration\footnote{One can also bound the velocity of the system. However, to simplify the exposition, we only constrain the system with acceleration bounds.}.

@@

\dy{Do you think we should include some figures that has been commented out below to show how difficult sampling an informed set is?}

@@

In this section we describe our approach to efficiently produce new samples in an informed set ~\Cinf given a specific cost~$c_{\text{best}}$ of trajectory $x_{best}(t)$.
The samples follow a Monte Carlo Markov Chain, in which a new sample candiate is produceded from a previous sample that also lies in the same informed set.
It is important to note that at each iteration, we will query our algorithm for one sample. 
However, this sample depends on the previous iterations of the algorithm which is why we describe it as a method to produce a sequence of samples.
Furthermore, the value~$c_{\text{best}}$ can decrease between consecutive iterations.
This will occur if the search algorithm that uses our sampler finds a path to the goal whose cost is lower than~$c_{\text{best}}$.

The idea behind introducing MCMC to informed sampling is to defining a target distribution $ \pi $ as a uniform distribution in a level set $ \Cinf $.
The sampling process consists of two stages,
\begin{enumerate}
	\item finding a sample in the informed same $ x_0 \in \Cinf $ as the start of a Markov chain, which is written as 
	$ x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( )$; and
	\item sampling a new sample $ x_i $ from a previous sample $ x_{i-1} $ in the same informed set, which is written as  $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$.
\end{enumerate}
Compared with rejection samplers, a MCMC-based informed sampler takes the advantage of a previous sample in the level set $ \Cinf $ in finding a new sample, while reserves the randomness in problem space search, which builds an ``exploitation and exploration" patter in informed sampling.

to initialize an MCMC algorithm which outputs a series of samples within \Cinf by constructing a Markov Chain.
This is done until some termination criterion is met and the entire process is repeated (Sec.~\ref{subsec:mcmc}). 
%Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}). 
For a visualization of the approach, see Fig.~\ref{fig:alg}.

\subsection{Find a sample in a level set}

In our setting, we are only interested in points in~\Cinf, thus we can start the Markov Chain in~\Cinf and avoid this burn-in stage. 
A new Markov chain is initialized when 
\begin{itemize}
	\item a new informed set is provided (e.g. a new current-best is found by the planner); or
	\item a new sample of an existing Markov chain is outside the level set.
\end{itemize}
There are a few methods that provides a sample in a given level set $ \Cinf $ in a planning problem, for example
\begin{itemize}
	\item sampling from the start state and the goal state
	\footnote{The start state and the goal state are guaranteed to be in any level set $ \Cinf $ because of the definition of a level set};
	\item using a gradient descent algorithm (e.g. Newton-Raphson Method~\cite{RT06}) to find a sample in the level set $ \Cinf $; 
	\footnote{ $ \forall x_i \in \Cinf \land x_j \not\in \Cinf, c( x_i ) > c ( x_j ) $. }
	\item sampling from a pool of previous samples that are in the level set $ \Cinf $; and
	\item calling a rejection sampling till one is in the level set is found.
\end{itemize}

Different methods have their pros and cons.
For example, a gradient descent algorithm is usually efficient in finding a solution, but subject to only convex problems.
Sampling from a pool of samples is algorithmic-free but might attract new samples near around previous samples.

In theory, MCMC methods converge to the desired distribution regardless of the initial sample used to seed the chain.
In our setting, the probability distribution~\Pinf is defined by having all points in \Cinf distributed uniformly
while 
the probability of sampling any configuration $x \in \calX \setminus \Cinf$ is zero.
A common practice to avoid starting biases in MCMC-type algorithm is to discard an initial set of samples (a process referred to as ``burn-in'')~\cite{ADDJ03}.

However, our cost function is not necessarily continuous (see Sec.~\ref{sec:mtdi}), thus if no sample was produced on $\partial \Cinf$ after a predefined number of steps, we restart the algorithm using a different random state~$x_\text{rand}$
(see also Sec.~\ref{subsec:mcmc}).

\subsection{Generate a new sample in a Markov chain}
\label{mcmc}

Our approach is general and can be applied to any MCMC algorithm (see Sec.~\ref{sec:related_work}).
We demonstrate how to instantiate it with two different algorithms: 
Metropolis-Hastings and Hit-and-Run.

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\LOOP
		%\STATE $x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( method )$
		%\WHILE {\texttt{MCMC\_continue\_chain()}}
		\STATE $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$ \label{alg:mcmc_informed_sampling:start}
		\IF{ $ x_{i} \not\in \Cinf $ }
			\STATE $x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( )$
			\STATE \textbf{Goto} line \ref{alg:mcmc_informed_sampling:start} 
		\ENDIF
		\STATE \textbf{output} $x_{i}$
		%\ENDWHILE
		\ENDLOOP 
		%		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
		%%		\STATE $x_{\text{boundary}} 
		%		\STATE $x_{0} 
		%		 \leftarrow
		%		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
		%%		\STATE $x_{0} \leftarrow x_{\text{boundary}}$
		%		\WHILE {\texttt{MCMC\_continue\_chain()}}
		%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$; \hspace{3mm} $i\leftarrow i+1$
		%			\STATE \textbf{output} $x_{i}$
		%		\ENDWHILE		
		%		\FOR {$i = 1$ to $N$}
		%			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$
		%			\STATE \textbf{output} $x_{i}$
		%		\ENDFOR
	\end{algorithmic}
	\caption{MCMC-based Informed Sampling $(c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

In MCMC-based Informed Sampling, as described in Algorithm~\ref{alg:mcmc_informed_sampling}, 
a sampler firstly 
\begin{enumerate}
\item generate a sample follows a Markov chain 
(\texttt{MCMC\_sample, line 2});
and then
\item decide if the candidate sample is in the level set, a new Markov chain will be created as a restart
(\texttt{sample\_in\_informed\_space, line 4}).
\end{enumerate}

\paragraph{Metropolis-Hastings sampler}

\subsubsection{Hit-and-Run sampler}
\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $d \leftarrow$ \texttt{sample\_random\_direction}$()$
		\STATE $ L(\lambda) = \{  x \mid x = x_{i-1} + \lambda d_i \} $
		\STATE $ \lambda^{+} \leftarrow \sup L(\lambda) $; 
				\hspace{3mm} 
			   $ \lambda^{-} \leftarrow \inf L(\lambda) $

@@

\ignore{
\begin{algorithm}[t]
	\begin{algorithmic}[1]
	\IF{ $method = \texttt{Newton-Rapson}$  }
		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
		\STATE $x_{0} \leftarrow \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$ ; \hspace{3mm} $i\leftarrow 0$
	\ELSIF{ $method = \texttt{Previous-Trajectory}$ }
	    \STATE $t_{rand} \leftarrow \texttt{sample\_random\_time()}$
	    \STATE $x_{0} \leftarrow x_{best}(t_{rand})$
	\ENDIF
	\RETURN $x_0$
	\end{algorithmic}
	\caption{sample\_in\_informed\_space$(method, c_{best})$}
	\label{alg:sample_inf}	
\end{algorithm}
}

@@

\begin{figure}[t!]
	\centering
	\begin{subfigure}[b]{0.78\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/levelset}
		\caption{2DoF - 4D. TODO:Time per sample vs level set}
		\label{fig:sampling_efficiency:levelset}
	\end{subfigure}
	%\begin{subfigure}[b]{0.48\linewidth}
	%	\includegraphics[width=\linewidth]{fig/sampling_efficiency/levelset_6d}
	%	\caption{6DoF - 12D. TODO:Time per sample vs level set}
	%	\label{fig:sampling_efficiency:6d}
	%\end{subfigure}
	 \\
	\begin{subfigure}[b]{0.78\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{2DoF - 4D TODO: rescale x axis}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
% 	\begin{subfigure}[b]{0.32\textwidth}
% 		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_3d}
% 		\caption{3DoF - 6D}
% 		\label{fig:sampling_efficiency:3d}
% 	\end{subfigure}	
	\begin{subfigure}[b]{0.78\linewidth}
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{6DoF - 12D. TODO: rescale x axis}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Sampling efficiency of different sampling methods}
	\label{fig:sampling_efficiency}
\end{figure} 
vantage of sampling a near state.
Both the per-sampling time of HRS 
It is noticeable that both x axis and y axis are converted by $ \log_{10} () $ in Figure~\ref{fig:sampling_efficiency:2d} and Figure~\ref{fig:sampling_efficiency:6d}.
This reserves the linearity of the data from rejection sampler, but also scales the difference in a smaller range for comparison.
Keep in mind that the differences in unit (ms) are much bigger.
\dy{TODO: change MCMC to MH (Metropolis Hasting) in the plots}
Metropolis-Hasting shows consistent performance when the informed set volume ratio is decreased, because it takes the adand HNR increases with the decrease of the informed set volume ratio. 
When informed set volume ratio is relatively high, it is easy to generate qualified samples.
All the samplers have close performances.
RS and HRS can be the most efficient because of its simple design.
But when problems get harder, HNR outperforms HRS significantly.

@@

\section{Conclusion}
\label{sec:future}

In this work we demonstrated the effectiveness of using MCMC algorithms to 
efficiently produce samples for asymptotically-optimal motion planning algorithms.
Clearly, there are multiple other MCMC algorithms that can be used and it is interesting to see if alternative algorithms may produce better results.
One drawback of these approaches is that they usually incur parameters that have to be tuned. Indeed, in this work we did not spend effort in tuning the parameters and did not change them across the range of scenarios we tested. 
There is a wealth of literature in the optimization community regarding this topic and integrating  such tools is left for future work.
Finally, we are interested in using this framework with alternative sampling-based algorithms such as BIT*~\cite{GSB15} or LBT-RRT~\cite{SH16} and with alternative state spaces.
%Here, a potential challenge lies with the fact the distance function we considered is not a metric, which may require changes
